---
title: "Project 1"
---

```{r global_options, include=FALSE}
library(knitr)
library(ggplot2)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)
```

*Enter your name and EID here*

**This project is due on the deadline posted on edX. Please submit as a pdf file.**


**Part 1:** Demonstrate basic command of Markdown by creating a bulleted list with three items, a numbered list with three items, and a sentence that has one word in bold and one word in italics.

---

#### Items
* Item 1
* Item 2
* Item 3

#### Things
1. Thing 1
2. Thing 2
3. Thing 3

Quick _brown_ fox __jumps__ over the **lazy** *dog*.

---

**Part 2:** The `mtcars` dataset contains information on car design and performance in 1974:

```{r}
mtcars
```

Use ggplot to make a line plot of the weight of a car (column `wt`) versus its fuel economy, measured in miles/gallon (column `mpg`).

```{r}
ggplot(mtcars, aes(wt, mpg)) + geom_line()
```

Now make a scatter plot (using `geom_point()`) of the weight of the car versus its horse power (`hp`), and color points by its fuel economy.

```{r}
ggplot(mtcars, aes(wt, hp, color = mpg)) + geom_point()
```
