---
title: "Homework 5"
---

```{r global_options, include=FALSE}
library(knitr)
library(ggplot2)
library(colorspace)
library(tidyverse)
library(broom)
library(sf)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)
```

**This homework is due on the deadline posted on edX. Please submit a .pdf file of your output and upload a .zip file with your .Rmd file.**

**Problem 1:** Use the color picker app from the **colorspace** package (`colorspace::choose_color()`) to create a qualitative color scale containing five colors. One of the five colors should be `#5C9E76`, so you need to find four additional colors that go with this one.

```{r fig.width = 6, fig.height = 3}
# replace "#FFFFFF" with your own colors
colors <- c("#5C9E76", "#FFFFFF", "#FFFFFF", "#FFFFFF", "#FFFFFF")
swatchplot(colors)
```


For the rest of this homework, we will be working with the `midwest_clean` dataset, which is a cleaned up version of the **ggplot2** `midwest` dataset.

```{r}
midwest_clean <- midwest %>% 
  select(
    state, county, area, popdensity, percbelowpoverty, inmetro
  ) %>%        # keep only a subset of data
  na.omit()    # remove any rows with missing data
```


**Problem 2:** Perform a PCA of the `midwest_clean` dataset and make a rotation plot of components 1 and 2.

```{r}
# your code goes here
```


**Problem 3:** Make a scatter plot of PC 2 versus PC 1 and color by state. You should use the custom colorscale you created in Problem 1. Then use the rotation plot from Problem 2 to describe where Chicago, Illinois can be found on the scatter plot. Provide any additional evidence used to support your answer.

```{r}
# your code goes here
```

*Your answer goes here.*



